---
title: 'sd_of_measures'
author: "[Orhan Aktaş](https://www.linkedin.com/in/orhanaktas/)"
date: "10 December 2020"
output:
  html_document:
    theme: flatly
    df_print: paged
  pdf_document: default
  word_document: default
   
---
# Introduction

The purpose of this analysis is to investigate the standart deviations of survey results for each measure. A high sd might signal a debate on a measure and this might be a valuable insight from data.


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r libraries}
library(googlesheets4)
library(tidyverse)
library(reshape2)
library(plotly)
library(psych)
```

Lets import data from google speadsheet and take a look at data.
```{r get_data, message=FALSE, warning=FALSE}
gs4_deauth()
link<-"https://docs.google.com/spreadsheets/d/1sXxCzr_EadTeXEK_KSMn5WMHI3UkAvF101pRmfK1yMA/edit?usp=sharing"
data<-read_sheet(link)

head(data)
```

Then it is good to see the sd's in a bar chart. A horizontal bar chart ordered by decreasing sd's is useful to investigate measures. 
```{r}
describe(data[3:81]) %>% 
  mutate(measure=colnames(data[3:81])) %>% 
  ggplot()+
  geom_col(mapping=aes(x=reorder(measure,sd),y=sd))+
  theme(axis.text=element_text(size=5))+
  coord_flip()
```


# Controversial Measures

Just analyze only the measures that has sd over 1. First see them in a boxplot.

```{r}

controversial_measures<-describe(data[3:81]) %>% 
  arrange(desc(sd)) %>%  
  filter(sd>1) %>% row.names()

data[2:81] %>% 
  pivot_longer(!Kurumunuz, names_to = "measure", values_to = "count") %>% 
  filter(measure %in% controversial_measures) %>% 
  ggplot()+
  geom_boxplot(
    mapping = aes(measure,count,fill=measure),
    show.legend = FALSE,
    outlier.shape = 1)+
  theme(axis.text=element_text(size=5))+ 
  scale_y_continuous(breaks=seq(1, 5, 1),limits = c(1,5)) +
  coord_flip()
```

Since boxplot doesn't mean much I prefer scatter plot with jittering.
```{r}

data[2:81] %>% 
  pivot_longer(!Kurumunuz, names_to = "measure", values_to = "count") %>% 
  filter(measure %in% controversial_measures) %>% 
  ggplot()+
  geom_jitter(mapping = aes(measure,count),width = 0.1,height = 0.1)+
  theme(axis.text=element_text(size=5))+ 
  scale_y_continuous(breaks=seq(1, 5, 1),limits = c(1,5)) +
  coord_flip()
```

It's very insightfull. You can see all the answers and their distrubition for each measure immediately just in a plot. 

# Other Measures
To understand how the answers differ, let's take a look at the situation with the measures with less bias.

```{r}

undisputed_measures<-describe(data[3:81]) %>% 
  arrange(desc(sd)) %>%  
  filter(sd<0.6) %>% row.names()

data[2:81] %>% 
  pivot_longer(!Kurumunuz, names_to = "measure", values_to = "count") %>% 
  filter(measure %in% undisputed_measures) %>% 
  ggplot()+
  geom_jitter(mapping = aes(measure,count),width = 0.1,height = 0.1)+
  theme(axis.text=element_text(size=5))+
  scale_y_continuous(breaks=seq(1, 5, 1),limits = c(1,5)) +
  coord_flip()

```

Finally I'll look at the measures that has average sd's.
```{r}

other_measures<-describe(data[3:81]) %>% 
  arrange(desc(sd)) %>%  
  filter(sd>0.6 & sd<1) %>% row.names()

data[2:81] %>% 
  pivot_longer(!Kurumunuz, names_to = "measure", values_to = "count") %>% 
  filter(measure %in% other_measures) %>% 
  ggplot()+
  geom_jitter(mapping = aes(measure,count),width = 0.1,height = 0.1)+
  theme(axis.text=element_text(size=5))+
  scale_y_continuous(breaks=seq(1, 5, 1),limits = c(1,5)) +
  coord_flip()

```

# Relation between sd and mean
After seeing all the distributions I have a doubt on the effect of last scale on sd. Now I'm looking to relation between mean and sd.
```{r}
p<-describe(data[3:81]) %>% 
  mutate(measure=row.names(describe(data[3:81]))) %>% 
  ggplot()+
  geom_point(mapping = aes(x=mean,y=sd,group=measure))+
  scale_x_continuous(breaks=seq(1, 5, 1),limits = c(1,5)) 
fig<-ggplotly(p)
fig
````
 As I doubted there is a negative correlation between mean and sd. When people think that one measure has importance they score 4 or 5 while in other situations their scores differ more. 





