---
title: 'Survey Results Factor Analysis'
author: "[Orhan Aktaş](https://www.linkedin.com/in/orhanaktas/)"
date: "10 December 2020"
output:
  html_document:
    theme: flatly
    df_print: paged
  pdf_document: default
  word_document: default
   
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  #fig.width=12, 
  #fig.height=8, 
  #fig.path='Figs/',
  echo=FALSE, 
  message=FALSE, 
  warning=FALSE)
```

```{r libraries}
library(googlesheets4)
library(tidyverse)
library(reshape2)
library(heatmaply)
library(psych)
library(magrittr)
<<<<<<< HEAD
=======
library(formattable)
>>>>>>> a9d1ccd101f70d695f068677e9903d72540f2a4a
library(DT)
```


```{r get_data, message=FALSE, warning=FALSE}
gs4_deauth()
link<-"https://docs.google.com/spreadsheets/d/1sXxCzr_EadTeXEK_KSMn5WMHI3UkAvF101pRmfK1yMA/edit?usp=sharing"
data<-read_sheet(link)
head(data)
```

# Sample Size and Sample to Variable Ratio
Sample size is `r nrow(data)` and there are `r ncol(data)-2` number of measures in the data set. 



create correlation matrix
```{r}
matrix<-cor(data[3:81],data[3:81])
```

histogram of correlations
```{r}
matrix %>% 
  melt %>% 
  mutate(value=abs(value)) %>% 
  filter(value<0.99) %>% 
  ggplot (mapping=aes(x=value)) + 
  geom_histogram(binwidth = 0.1,boundary =1)+ 
  scale_x_continuous(name="correlation", breaks=seq(0, 1, 0.1),limits = c(0,1))
```

# Is the data suitable for factor analysis?

the measures of factoring adequacy
or
whether the matrix is factorable

[ref:](https://cran.r-project.org/web/packages/psych/psych.pdf)

Bartlett test of Sphericity which tests that a correlation matrix is an identity matrix
```{r}
cortest.bartlett(matrix)
```

Kaiser, Meyer, Olkin Measure of Sampling Adequacy
```{r}
KMO(matrix)
```


determinant is zero so I tried to exclude some variables but it is still 0 after excluding 18 variables.
```{r}
# variables with most correlations over 0.7
delete2<-matrix %>% 
  melt %>%
  filter(value<0.99 & value>0.7) %>% 
  group_by(Var1) %>% 
  summarise(n=n()) %>% 
  arrange(desc(n)) %>% 
  head(18) %>% 
  extract(1) %>% 
  t %>% 
  as.character

data2<-data %>% select(-any_of(delete2)) 
matrix<-cor(data2[3:ncol(data2)],data2[3:ncol(data2)])
det(matrix)
```

factor analysis for x variables that has minimum average correlations
```{r}
# variables with most correlations over 0.7
delete2<-matrix %>% 
  melt %>%
  filter(value<0.20) %>% 
  group_by(Var1) %>% 
  summarise(mean=mean(abs(value))) %>% 
  arrange(mean) %>% 
  head(20) %>% 
  extract(1) %>% 
  t %>% 
  as.character

data2<-data %>% select(all_of(delete2)) 
matrix2<-cor(data2[1:ncol(data2)],data2[1:ncol(data2)])
det(matrix2)
#KMO(matrix2) 25 taneye sonuç veriyor.
VSS.scree(matrix2, main = "scree plot")
results<-factanal(data2, 6, rotation="varimax")
table<-results$loadings[,1:6] %>% as_tibble() %>% mutate(measure=results$loadings[,1:6] %>% row.names) %>%  select(measure, everything())
#table %>% write_excel_csv("results_6.csv")
table<-table %>% pivot_longer(!measure, names_to = "factor", values_to = "loading") %>% group_by(measure) %>% filter(loading==max(loading)) %>% select(measure,factor,loading) %>% arrange(factor,desc(loading))
table
```

<<<<<<< HEAD
# table format
=======
table format
```{r}

formattable(table, 
            align =c("l","c","c"), 
            list(`factor`= formatter("span",style = x ~ style(
              font.weight = "bold", 
              color = 	ifelse(x =="Factor1", "red",
			      	            ifelse(x== "Factor2", "green",
				                    ifelse(x== "Factor3", "blue",
				                      ifelse(x== "Factor4", "yellow",
				                        ifelse(x== "Factor5","orange",
				                          ifelse(x== "Factor6", "purple", "black")
				                        )
				                      )
				                    )
			      	            )
                  )
                )
              )
            )
)
```
>>>>>>> a9d1ccd101f70d695f068677e9903d72540f2a4a

with DT formatStyle function in DT package it is possible to change all cells in a row base on a value from a cell in that row. 
```{r}
table %>% 
	datatable %>% 
	formatStyle( 'factor',
	       fontWeight = "bold",
		     target = 'row',
		     color = styleEqual(
					c("Factor1","Factor2","Factor3","Factor4","Factor5","Factor6"),
<<<<<<< HEAD
					c("red","green","#b8860b","blue","orange","purple")
=======
					c("red","green","blue","yellow","orange","purple")
>>>>>>> a9d1ccd101f70d695f068677e9903d72540f2a4a
				      )
	) %>% 
  formatPercentage('loading', 0)
```


heatmap
```{r eval=FALSE, warning=FALSE, include=FALSE}


#png("high_res.png", width = 8, height = 8, units = 'in', res = 1600)
#heatmap(as.matrix(data[3:ncol(data)]))

heatmaply_cor(
  matrix, 
  showticklabels = c(F,F), 
  margins = c(80,10),
  label_format_fun = function(...) round(..., digits = 1))
#dev.off()
```

